{\bf Practical Concurrency Testing} \\ \normalsize \vspace{1em} \begin{tabular}{c} {\em or: How I Learned to Stop Worrying and Love the Exponential Explosion} \end{tabular}

### 2018-12-24 thesis/intro.tex
assorted minor proofreads

Pronoun use will vary between more specific and more ambiguous to convey additional nuance.
The singular ``I'' is associated with my own research contributions,
while the royal ``we'' should be taken to include the reader,
such as when surveying background material or related work,
to which the author and reader share more similar relationships.
The impersonal ``the programmer'' will be referred to as she/her
to highlight her role as the intended user, separate from the underlying research,
as well as to challenge readers' unconscious bias about gender in computer science.
Gender-neutral pronouns will be used to refer to
individual students who participated in the user studies,
%will be given the more inclusive they/them.
%Gender-neutral pronouns will also be used
as well as
on the author themself,
who would also prefer them for any third-person reference by you.

>>>>

Pronoun use will vary between more specific and more ambiguous to convey additional nuance.
The singular ``I''
%is associated with
connotes
my own research contributions,
while the royal ``we'' should be taken to include the reader,
such as when surveying background material or related work,
to which the author and reader share more similar relationships.
The impersonal ``the programmer'' will be referred to as she/her
to highlight her role as the intended user, separate from the underlying research,
as well as to challenge readers' unconscious bias about gender in computer science.
Gender-neutral pronouns will be used to refer to
individual students who participated in the user studies,
%will be given the more inclusive they/them.
%Gender-neutral pronouns will also be used
as well as
on the author themself.
%who would also prefer them for any third-person reference by you.

### 2018-12-24 thesis/landslide.tex
assorted minor proofreads

		\footnote{These two options replace the deprecated {\tt kern\_ready\_for\_timer\_interrupt()} annotation from {\tt student.c} described in \cite{landslide} \sect{6.2.3}.}

@@

The symbol table logic lives in {\tt symtable.c} and is pretty much a lot of glue code.
In the Simics version, Landslide relies on the {\tt deflsym} Simics object created by the 15-410 python scripts,
and queries its attributes using Simics API calls.
In the Bochs version, function names and line numbers are handled separately:
Bochs is patched with a new API function named {\tt bx\_dbg\_symbolic\_address\_landslide}%
\footnote{does the same thing as the existing {\tt bx\_dbg\_symbolic\_address}, but with a better type signature}
which provides function names and hexadecimal offsets;
while for line numbers, {\tt pebsim/pintos/build.sh}%
\footnote{Line numbers in Bochs for Pebbles/P2s are not supported yet; see {\tt p2-setup.sh} for the work-in-progress.}
generates a header file {\tt line\_numbers.h}
using {\tt objdump} and {\tt addr2line},
which the aforementioned hex offset then serves as an index into.

@@

The stack tracer is implemented in {\tt stack.c}.
It does the standard approach of following the base pointer chain
(not supporting code compiled with {\tt -fomit-frame-pointer} by doing anything clever
like understanding how much stack frame is allocated for each function),
and printing symbol table information for the pushed return addresses at the top of each frame.

However, it also offers several special-case features
which even some students have sometimes noticed as being more clever than Simics's stack tracer.
I document those features here.
As a common point of implementation among them,
Landslide traces the stack pointer {\tt esp} in addition to the base pointer {\tt ebp};
not only updating it whenever dereferencing the base pointer,
but also when decoding simple assembly routines, finding ``hidden'' stack frames without base pointers,
identifying system call wrappers, and so on.
The corresponding code lives in {\tt stack\_trace()} in {\tt stack.c}.

@@

\begin{itemize}
	\item If no {\tt within\_function} commands are given, operate in ``denylist mode'':
		the preemption point is by default valid as long as no {\tt without\_function} calls reject it.
		Otherwise, operate in ``allowlist mode'':
		the preemption point is by default rejected unless at least one {\tt within\_function} directive matches.
	\item Subject to the above, find the sequentially-last {\tt *\_function} directive
		(static preemption point commands considered before dynamic ones)
		which matches any function in the stack trace.
		If {\tt within}, accept the preemption point; if {\tt without}, reject it.
\end{itemize}

@@

\subsubsection{Run-time (Recursive) estimation}

>>>>

		\footnote{These two options replace the deprecated {\tt kern\_ready\_for\_timer\_interrupt()} annotation from {\tt student.c} described in \cite[\sect{6.2.3}]{landslide}.}

@@

The symbol table logic lives in {\tt symtable.c} and is pretty much a lot of glue code.
In the Simics version, Landslide relies on the {\tt deflsym} Simics object created by the 15-410 python scripts,
and queries its attributes using Simics API calls.
In the Bochs version, function names and line numbers are handled separately:
Bochs is patched with a new API function named {\tt bx\_dbg\_symbolic\_address\_landslide()}%
\footnote{Does the same thing as the existing {\tt bx\_dbg\_symbolic\_address()}, but with a better type signature.}
which provides function names and hexadecimal offsets;
while for line numbers, {\tt pebsim/build.sh} (and {\tt pebsim/pintos/\allowbreak{}build.sh} for Pintos)
%\footnote{Line numbers in Bochs for Pebbles/P2s are not supported yet; see {\tt p2-setup.sh} for the work-in-progress.}
generates a header file {\tt line\_numbers.h}
using {\tt objdump} and {\tt addr2line},
which the aforementioned hex offset then serves as an index into.

@@

The stack tracer is implemented in {\tt stack.c}
using the standard approach of following the base pointer chain
(not supporting code compiled with {\tt -fomit-frame-pointer} by doing anything clever
like understanding how much stack frame each function allocates),
and printing symbol table information for the pushed return addresses at the top of each frame.
%
However, it also offers several special-case features
which even some students have sometimes noticed as being more clever than Simics's stack tracer.
I document those features here.
As a common point of implementation among them,
Landslide traces the stack pointer {\tt esp} in addition to the base pointer {\tt ebp};
not only updating it whenever dereferencing the base pointer,
but also when decoding simple assembly routines, finding ``hidden'' stack frames without base pointers,
identifying system call wrappers, and so on.
The corresponding code lives in {\tt stack\_trace()} in {\tt stack.c}.

@@

\begin{itemize}
	\item If no {\tt within\_function} commands are given, operate in ``denylist'' mode:
		the preemption point is by default valid as long as no {\tt without\_function} calls reject it.
		Otherwise, operate in ``allowlist'' mode:
		the preemption point is by default rejected unless at least one {\tt within\_function} directive matches.
	\item Subject to the above, find the sequentially-last {\tt *\_function} directive
		(static preemption point commands considered before dynamic ones)
		which matches any function in the stack trace.
		If {\tt within}, accept the preemption point; if {\tt without}, reject it.
\end{itemize}

@@

\subsubsection{Runtime (Recursive) estimation}

### 2018-12-24 thesis/quicksand.tex
assorted minor proofreads

\qrevision{When configuring a model checker's preemption strategy,
or indeed,
choosing a model checker to begin with,
the resulting state space is {\em parameterized} by the set of preemption points.
%
In the first example of \Cref{fig:tree}
I constructed the state space by expanding both threads' {\tt x++} operations into three pseudo-assembly instructions,
then designating every instruction as a possible preemption point,
yielding ${3+3 \choose 3} = 20$ total interleavings.
Later, \cref{sec:landslide-dpor} showed DPOR prune 16 of those as equivalent,
although the reduced state space size is still combinatorial
in the number of {\em conflicting} events, rather than the total number.
For larger tests, %with hundreds or thousands of conflicting events in any given execution,
committing in advance to test all possible interleavings quickly becomes impractical.
Accordingly, many existing model checkers opt for preempting on only a subset of execution events,
such as synchronization API boundaries.

>>>>

\qrevision{When configuring a model checker's preemption strategy,
or indeed,
choosing a model checker to begin with,
the resulting state space is {\em parameterized} by the set of preemption points.
%
In the first example of \Cref{fig:tree}
I constructed the state space by expanding both threads' {\tt x++} operations into three pseudo-assembly instructions,
then designating every instruction as a possible preemption point,
yielding ${3+3 \choose 3} = 20$ total interleavings.
Later, \cref{sec:landslide-dpor} showed that DPOR prunes 16 of those as equivalent,
although the reduced state space size is still combinatorial
in the number of {\em conflicting} events, rather than the total number.
For larger tests, %with hundreds or thousands of conflicting events in any given execution,
committing in advance to test all possible interleavings quickly becomes impractical.
Accordingly, many existing model checkers opt for preempting on only a subset of execution events,
such as synchronization API boundaries.
